---
title: "Exploratory Data Analysis of the Red Wine dataset"
author: "Ravinder Bahra"
date: "April 15th 2018"
output: 
  html_document:
    css: eda.css
---
___


```{r echo=FALSE, clean_workspace}
# clear the workspace
rm(list=ls())

```


```{r echo=FALSE, install_dependencies, warning=FALSE, message=FALSE}
# install all dependencies if not installed
# install.packages("ggplot2", dependencies = T, repos = "http://cran.us.r-project.org") 
# install.packages("knitr", dependencies = T, repos = "http://cran.us.r-project.org")
# install.packages("dplyr", dependencies = T, repos = "http://cran.us.r-project.org")
# install.packages("GGally", dependencies = T, repos = "http://cran.us.r-project.org")
# install.packages("lemon", dependencies = T, repos = "http://cran.us.r-project.org")
```

```{r echo=FALSE, load_dependencies }
suppressWarnings(suppressMessages(library(ggplot2)))
suppressWarnings(suppressMessages(library(gridExtra)))
suppressWarnings(suppressMessages(library(GGally)))
suppressWarnings(suppressMessages(library(dplyr)))
# This is used for pretty printing data frame tables
suppressWarnings(suppressMessages(library(lemon)))
suppressWarnings(suppressMessages(library(reshape)))

knit_print.data.frame <- lemon_print
```

```{r echo=FALSE, load_dataset}
# Read 
df <- read.csv('./datasets/wineQualityReds.csv')
# df <- read.csv('D:/RFiles/datasets/wineQualityReds.csv')
wine_dataset <- df[,2:length(df)]
```

## Introduction

The following report uses the public red wine dataset collected for the report
titled *Modeling wine preferences by data mining from physicochemical properties
*by Paulo Cortez.

The dataset consists of 1599 samples, 11 input variables which are based on 
physiochemical properties of the wine and 1 output variable which is the quality of the wine.
The quality of the wine was based on sensory data (taste).

3 wine experts assessed and rated the quality of the wine, a score between 
0 to 10 was given where 0 is very bad and 10 is very good.

This report will investigate which of these properties impacts the quality of 
the wine.

___

## Univariate Plots

### View of the data  {.smallfont}

```{r echo=FALSE, head_wine_dataset}
head(wine_dataset)
```



## {}

#### Statistical summaries of wine properties
```{r echo=FALSE, summary_wine_dataset}
summary(wine_dataset)
```

The dataset loads ok from the csv file, we can see the 12 variables which are
all numerical.

You can get a sense of the distributions from having a quick look of summary
table. We'd get a better look from viewing the distributions as histograms which
is shown below

#### Distributions of wine properties

The following code will loop over the column names and generate histograms of
all the wine properties
```{r echo=FALSE, distribution_wine_properties}
# this code will iterate over all columns/properties in the dataset and create
# a distribution. The distributions wll be added to a list. This list will be
# passed to grid.arrange to arrange a grid view of all the distribution plots

# pull all the names of the columns into a vector variable
column_names <- names(wine_dataset)

# create an empty list to store the plots 
plot_list <- list()

# iterate over the column names and create a plot for each property
for (i in 1:length(column_names)){
  # generate a histogram and add to the list
  plot_list[[i]] <- ggplot(aes_string(x=column_names[i]), data=wine_dataset) +
    geom_histogram(bins=30)
}
# pass the list to grid.arrange to display all pl
grid.arrange(grobs = plot_list, ncol = floor(sqrt(length(column_names))))
```

Both density and pH stand out as normally distributed, other distributions look
positively skewed. 

#### Counts for the variable Quality of the wines

```{r echo=FALSE, table_counts_quality}
table(wine_dataset$quality)
```

Quality is an ordinal variable from 1-10 which was the rating given by the wine
taster. We can see from the table above most wines were rated 5-6. No wines were
given a rating of 1,2,9 or 10. 

Not all the plots in the grid are easily visible from the above table, I will look at residual
sugar and chlorides below again but larger.

#### Residual Sugar 

The below plot shows the residual sugar applied with a logarithmic scale. The
original was positively skewed, the log scale was applied to make the
distribution look more normal

```{r echo=FALSE, dist_residual_sugar}
qplot(x=residual.sugar, data=wine_dataset, bins=30) +
  scale_x_log10()
```

#### Chlorides 

```{r echo=FALSE, dist_chlorides}
qplot(x=chlorides, data=wine_dataset, bins=30) +
  scale_x_log10()
```

## Univariate Analysis


### What is the structure of the dataset?

The dataset contains `r nrow(wine_dataset)` rows and 12 columns, 11 of those
columns are measured properties of the wine and are continuous variables. The
variable Quality is an ordinal variable. Quality is a score from 1-10 given
by the wine taster. Most of the wines were given a rating of 5-6 and none of the
wines in the dataset were rated 1,2, 9 or 10.

The table shows that all of the data is numerical, we have the following 
physicochemical properties we can use to explore the dataset. Some of the explanations
for the properites were given as part of the dataset.

- Fixed Acidity - Represents the non volatile acids in the wine
- Volatile Acidity - Too high can results in a vinegar like taste (unpleasant)
- Citric Acid - adds flavour to the wine
- Residual Sugar - The sugar left after the fementation process. Large amounts 
are found in sweeter wines
- Chlorides - salt content
- Free Sulfur Dioxide  - used to prevent microbial growth and prevents the wine
from oxidizing
- Density - density of the liquid. Denisity of water is  0.997 g/m<sup>3</sup>
- pH - The acidity of the wine, lower values on the ph scale are more acidic (higher is alkaline). According to the information given with the dataset we'll find most wines have a ph between 3-4
- Sulphates - added to the wine to increase the sulfur dioxide gas level. Sulfur
dioxide gas is an antimicrobial (agent that kills microorganisms )
- Alcohol - The % of alcohol in the wine 


### What are the main features of interest in your dataset?

I'll be exploring how some of the physicochemical properties effect quality.
At this stage it is difficult to say which features are going to be the primary
focus, that will become clearer during the bivariate analysis. However i'm 
interested in seeing how diffierent quantities of citric acid changes the
perception of quality of the wine. Residual Sugar which makes the wine sweet may
also have an effect.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

Other properties of acidity and alcohol are also of interest.

### Other observations

Looking at the distribution of the wine quality, most of the wines were rated
5/10 - 6/10. The lowest rating was actually 3 and the highest was 8. 

Most of the distributions are positively skewed, density, ph and quality appear
normally distributed. 

Citric acid shows a bimodal distribution.

Looking at the summary statistics for each physicochemical property, some of
the propeties have a lot of outliers. The most notiable from the distributions
and the summary statistics are 

- residual sugar
- chlorides
- total sulfur dioxide
- sulphates

In the bivariate section we can begin to look at how these variables
effect each other to determine what properties effect the quality.

___

# Bivariate Plots

Now we will look at multiple variables together. Pairwise plots are a good
way to see what variables correlate together

```{r echo=FALSE}
# The function scatter_func returns a scatter plot with the addition of a 
# regression line. The default behavior is over ridden by passing a function
# as a parameter in ggpairs

scatter_func <- function(data, mapping,...){
  p <- ggplot(data=data, mapping = mapping) +
    # add alpha level to the plot
    geom_point(alpha=0.01) +
    # add a regression line
    geom_smooth(method = "lm", se = FALSE) +
    #reduce the size of the text so it doesn't overlap
    theme(text = element_text(size=8),
        axis.text.x = element_text(angle=90))
}

colnames <- names(wine_dataset)[1:length(names(wine_dataset))]

ggpairs(wine_dataset,
        columns = 1:ncol(wine_dataset),
        upper = list(continuous = wrap("cor", size = 2.5)),
        lower = list(continuous = scatter_func),
        columnLabels = gsub('.', ' ', colnames, fixed = T),
        labeller = label_wrap_gen(10)
        #diag = list(continuous = diag_func)
        )
```

#### A look into Which variables correlate with quality

The output variable (quality) is of most interest to us, we'd like to determine 
which properties of the wine contribute to a greater quality

There are some properties that do correlate with quality. The correlations are
not strong but do stand out and are worth looking into for further analysis.

The properties that stood out against quality the most are as follows

physicochemical properties that correlate with quality positively are

- Alcohol content (+0.476)
- sulphate (+0.251)
- Citric acid (+0.226)

physicochemical properties that correlate with quality negatively are

- Volatile acidity (-3.91)

Other correlations not against quality are as follows

- *Total Sulfur Dioxide* and *Free Sulphur Dioxide* corrlate, which isn't unexpected
as total sulfur dioxide contains free sulfur dioxide
- *Citric Acid* and *Fixed Acidity* (0.672) - again this is expected with larger 
volumes of acid
- *pH* with *Citric Acid* (-0.542) and *pH* with *Fixed Acidity* (-0.683) - again will
correlate, higher acidity has a lower PH (higher Ph is more alkaline)
- *Citric Acid* and *Volatile Acidity* (-0.552)
- *Alcohol Content* and *Density* (-0.496)
- *Density* and *Residual Sugar* (0.355) 
- *Density* and *pH* (-0.342)

We can also see the correlations with quality as a table

## {.smallfont}

```{r echo=FALSE,warning=FALSE, correlation_data}
# correlate the columns in the wine dataset and add to a data table
df_corr <- data.frame(wine_properties = names(wine_dataset)[1:11], correlation = cor(x=wine_dataset[,1:11], y=wine_dataset$quality)[1:11])

# sort the table, this doesn't exactly sort the table in order of correlation strenght, as you h
df_corr_sorted <- arrange(df_corr, correlation)
df_corr_sorted
```

## {}


Based on the plots, i'd like to invesgiate the main 4 that correlated with quality

- Alcohol Content
- Sulphate
- Citric Acid
- Volatile Acidity

I'm also interested in the properties that showed a correlations amongst 
other non quality properties, such as 

- citric acid
- Density and Alcohol
- Density and residual sugar
- Density and pH


#### Box plots of quality

Let's take a look at some of the physicochemical properties against quality

```{r echo=FALSE}
# create a new variable to hold factor version of quality
wine_dataset$quality_factors <- as.factor(wine_dataset$quality)

# create box plots for each property

# alcohol
b_plot_alcohol <- ggplot(aes(x=quality_factors,y=alcohol),
      data=wine_dataset) +
  geom_boxplot(outlier.colour="red")

# volatile acidity
b_plot_acidity <- ggplot(aes(x=quality_factors,y=volatile.acidity),
      data=wine_dataset) +
  geom_boxplot(outlier.colour="red")

# citric acid
b_plot_citric_acid <- ggplot(aes(x=quality_factors,y=citric.acid),
      data=wine_dataset) +
  geom_boxplot(outlier.colour="red")

# sulphates
b_plot_sulphate<- ggplot(aes(x=quality_factors,y=sulphates),
      data=wine_dataset) +
  geom_boxplot(outlier.colour="red")

# arrange plots in a grid
grid.arrange(b_plot_alcohol,b_plot_acidity,b_plot_citric_acid, b_plot_sulphate)


```

- An increase in Alcohol seems to show an increase in quality.
- Volatile acidity content is less for better wines.
- Higher citric acid is seen in better quality wine
- Higher volume of suplhate is seen in better quality wine

```{r echo=FALSE, scatter_functions, message=FALSE}
# the following code allows us to pass in a property and it will create a 
# scatter plot
quality_scatter <- function(wine_property){
  plt <- ggplot(aes_string(x='quality',y=wine_property),data=wine_dataset) +
  # geom_point(position='jitter', aes(color=quality_factors), alpha=0.3) +
  geom_point(position='jitter', alpha=0.3) +
  scale_x_continuous(breaks=seq(2,9,1)) +
  geom_smooth(method = 'lm')
  return(plt)
}

# the function creates faceted historgrams which are split by the quality of the
# wine
wine_property_hist <- function(wine_property){
  plt <- ggplot(aes_string(x=wine_property), data=wine_dataset) +
  geom_histogram() +
  #geom_histogram(aes(fill=quality_factors)) +
  facet_wrap(~quality_factors)
  return(plt)
}
```

```{r echo=FALSE, create_factors, message=FALSE}
q_factors <- cut(df$quality, c(2,4,6,8))
# i've separated the wines into 3 groups of bad, ok and great wine, while
# factors were created earlier, it makes sense for the distributions to have
# fewer colors to not overload the charts, it also gives you a better sense of the relationship

wine_dataset$q_factors <- factor(q_factors, labels = c("bad", "ok", "great"))
```

#### Alcohol vs Quality

```{r echo=FALSE,message=FALSE}

quality_scatter('alcohol')


wine_property_hist('alcohol') + 
  scale_x_log10()

qplot(x=alcohol, data=wine_dataset, fill=q_factors) +
  scale_x_log10()
```

Using a log scale did improve the alcohol distribution, while still not normal
it is still less skewed than before.

#### Volatile Acidity vs Quality

```{r echo=FALSE, message=FALSE}

quality_scatter('volatile.acidity')

wine_property_hist('volatile.acidity') + scale_x_sqrt()

qplot(x=volatile.acidity, data=wine_dataset, fill=q_factors) + scale_x_sqrt()
```

A square root scale for volatile acidity improved the distribution which 
became more normal for ok and bad wines. Ok and bad wines have a similar 
distribution but great wines do not

#### Citric Acid vs Quality

```{r echo=FALSE, message=FALSE}

quality_scatter('citric.acid')

wine_property_hist('citric.acid') + 
  scale_x_sqrt()
  # scale_x_log10()

qplot(x=citric.acid, data=wine_dataset, fill=q_factors) +
  scale_x_sqrt()
  #scale_x_log10()
```

Citric acid shows a possible outlier on the low end of the spectrum

#### Sulphates vs Quality

```{r echo=FALSE, message=FALSE}

quality_scatter('sulphates')

wine_property_hist('sulphates') + 
  scale_x_log10()

qplot(x=sulphates, data=wine_dataset, fill=q_factors) +
  scale_x_log10()
```

Bad and ok wines show a similar distribution

#### Density

The density plot below shows the distribution of density with a line showing
where the density of water is.


```{r echo=FALSE, density_plot, message=FALSE}
ggplot(aes(x=density), data=wine_dataset) +
  geom_histogram() +
  geom_vline(xintercept=0.997, colour="red")
```

Let's take a look where the position of the distributions are when
using face wrap with alcohol factors.

```{r echo=FALSE, message=FALSE}
wine_dataset$alcohol_factors <- cut(wine_dataset$alcohol, breaks = seq(8.5,15,0.5))

ggplot(aes(x=density), data=wine_dataset) +
  geom_histogram() +
  geom_vline(xintercept=0.997, colour="red") +
  facet_wrap(~alcohol_factors)

```

We can see that as alcohol has increased the distribution of the density 
shifts towards the left, meaning the liquid becomes less dense the more alcohol
content there is in the liquid. 

## Bivariate Analysis 

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

Higher alcohol content is found in wines of better quality. Alcohol also effects
the density of the liquid, the more alcohol there was the less dense it became. 
This is because the density of alcohol is 0.789 g/m<sup>3</sup> compared to
water, which has a density of 0.997 g/m<sup>3</sup>.

Volatile acid is found much less in higher quality wines. According to the 
supplementary information found acetic acid, which is one of the primary
volatile acids is associated with the smell and taste of vingar. 

Citric acid increased with better quality wines. The supplementary information
informs us that citric acid adds freshness and flavour to wines. Citric acids 
is added after fermentation, it added before then the yeast can convert the acid
into acetic acid (Volatile). As we discoveres this is undesirable.

Higher sulphate was also seen in higher quality wines. Suphates are added to 
increase the suldur dioxide content which is an anti oxidant and antimicrobial.
Antimicrobials are important to prevent microorganisms from proliferating in the
wine. 

### What was the strongest relationship you found?

I found the quality of the wine increase with less acetic acid and more citric
acid.

___

## Multivariate Plots

This section will now explore more than 2 variables in the plots.

### Parallel coordinate plot
Parallel plots can show you general trends with all the variables mapped with 
parallel coordinate axes.

```{r echo=FALSE, message=FALSE}
# create a separate variable called
wine_props <- c("fixed.acidity",
               "volatile.acidity",
               "citric.acid",
               "residual.sugar",
               "chlorides",
               "free.sulfur.dioxide",
               "total.sulfur.dioxide",
               "density",
               "pH",
               "sulphates",
               "alcohol",
               "quality")

wine.original <- wine_dataset[wine_props]
#wine.original$quality <- as.factor(wine.original$quality)
ggparcoord(wine.original, groupColumn = 12, alphaLines = 0.1) +
  ggtitle("Parallel coordinates graph, darker blues are lower quality") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position="none",
        )
```

You can see some trends, such as darker colors (lower quality) have much less
citric acid, higher volatile acidity. We can see lighter colors (higher quality)
associated with higher sulphates and higher alcohol content.

I can also just look at the same chart but the mean of each property grouped by
quality

```{r echo=FALSE, message=FALSE}
# calculate the means of each property grouped by the quality
mean.table <- wine.original %>% group_by(quality) %>%
  summarise(
    m.fixed.acidity=mean(fixed.acidity),
    m.volatile.acidity=mean(volatile.acidity),
    m.citric.acid=mean(citric.acid),
    m.residual.sugar=mean(residual.sugar),
    m.chlorides=mean(chlorides),
    m.free.sulfur.dioxide=mean(free.sulfur.dioxide),
    m.total.sulfur.dioxide=mean(total.sulfur.dioxide),
    m.density=mean(density),
    m.pH=mean(pH),
    m.sulphates=mean(sulphates),
    m.alcohol=mean(alcohol)
  )

ggparcoord(mean.table, groupColumn = 1) +
  ggtitle("Parallel coordinates graph, darker blues are lower quality") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position="none",
        )

```

A few more trends become easier to visualise. Lower quality wines show less
sulphur dioxide, and higher quality wines show there is more SO2.Higher quality
wines are less dense and have more alcohol which supports my earlier discoveries

#### Sulfites

The sulfites are composed of both free sulfor dioxide and the bound form of
sulfur dioxide. The dataset contains the total sulfur dioxide but if we subtract
the free sulfur dioxide from the total sulfur dioxide, we will be left with
bound sulfur dioxide

I  want to look at the relationship between the free and bound forms of
sulfites. The total sulfur dioxide is a combination of free sulfur dioxide and 
bound sulfur dioxide. Lets see if there is a relationship between those 
variables

```{r echo=FALSE, bivariate_sulfites, message=FALSE}
# create a new variable for bound sulfur called bound.sulfur.dioxide
wine_dataset$bound.sulfur.dioxide <- with(wine_dataset, (total.sulfur.dioxide - free.sulfur.dioxide))

all_plts <- ggplot(aes(x=free.sulfur.dioxide, y=bound.sulfur.dioxide),data=wine_dataset) +
  geom_point(position='jitter', aes(color=quality_factors), alpha=0.2) +
  coord_trans(y='log', x='log')

fct_plts <- all_plts + facet_wrap(~quality_factors)

grid.arrange(all_plts, fct_plts, ncol=1)
  
```

The results dont look too promising. It's hard to tell with the graphs as there
are fewer results for the extrememly low and extremely high qualities.

Free form sulfur is the sulfur dioxide not bound to any other substances. Bound
has joined bonds with other substances which means it is no longer SO2. Free
Sulfir dioxide is what actually contributes to minimsing microorganisms and 
prevents oxidation. I expected to see an increase in bound sulfur and reduced
quality of wine. We can see a small relationship like this, most of the 8 rated
wines are in the bottom half of the chart where there is less bound sulfur.

This probably is not the best way to test this theory, it would be better to 
look at the proportion/ratio of free to bound and seeing a relationship.

We don't need the bound sulfur variable for this, we can easily construct a new
chart using a % of total sulfur dioxide.

This chart is not multivariate however i've kept it in this section to keep the
flow.

```{r echo=FALSE, message=FALSE}
wine_dataset$percentage.free.so2 <- ceiling((wine_dataset$free.sulfur.dioxide / 
                                       wine_dataset$total.sulfur.dioxide)*100)

ggplot(aes(x=percentage.free.so2, y=quality_factors), data=wine_dataset) +
  geom_point(position='jitter', aes(color=quality_factors), alpha=0.3)
```

There still doesn't seem to be any correlation.

#### Exploring Acid

I'd now like to investigate how quantities of acids effect the quality of the
wine, I'm going to look at volatile acidity against fixed acidity and citric
acid.

```{r echo=FALSE, message=FALSE}


volatile_v_citric <- ggplot(aes(x=volatile.acidity, y=citric.acid), data=wine_dataset) +
  geom_point(aes(color=q_factors), position='jitter', alpha=0.2) + 
  scale_color_brewer(type='qual', palette=1)


volatile_v_fixed <- ggplot(data = wine_dataset,aes(y = fixed.acidity, x = volatile.acidity)) +
  geom_point(aes(color=quality_factors), position='jitter', alpha=0.2)

grid.arrange(volatile_v_citric,volatile_v_fixed, ncol=2)


```

Looking at the comparison of volatile acidity against citric acid, we can see
that lower volatile acidity and higher citric acid contributes to a better
quality wine. All the low quality wines cluster in the region of high acetic
acid and low citric acid

As expected, ligher colors (higher quality) seems to be on the low end of the
volatile acid spectrum on the second chart. Fixed acidity however doesn't seem
to explain as much as volatile acid does.

#### Exploring sulphate

Sulphate is an additive which is used to increase the amount of sulfur dioxide
levels. I look at sulphates compared to both alcohol and chlorides.

```{r echo=FALSE, sulphates_alcohol, message=FALSE, warning=FALSE}
#sulphates vs alcohol
ggplot(aes(x=sulphates, y=alcohol), data=wine_dataset) +
  geom_point(aes(color=quality_factors), position='jitter', alpha=0.2)
```

We've explored alcohol extensivley so it's not surprising to see that more
alcohol increases quality. An increase in sulphates upto a point helps, which is
around 1. Beyond that it doesn't seem to help even with increased alcohol.

There also seems to be more sulphate in the wines with less alcohol.

```{r echo=FALSE,sulphates_chlorides, message=FALSE, warning=FALSE}

# need to use quantile as most of the data is clustered in a small area.
# Quantile of 99% will reduce the outliers in the graph
quantile.level <- 0.99

ggplot(aes(x=sulphates, y=chlorides), data=wine_dataset) +
  geom_point(aes(color=quality_factors), position='jitter', alpha=0.2) +
  xlim(0.25 ,quantile(wine_dataset$sulphates, quantile.level)) +
  ylim(0,quantile(wine_dataset$chlorides, quantile.level))
```

Most of the chlorides are clustered around 0.1, it doesn't seem to be effected
by the quantity of sulphates however.



#### Exploring Density

```{r echo=FALSE, message=FALSE}
ggplot(aes(x=density, y=alcohol), data=wine_dataset) +
  geom_point(aes(color=quality_factors), position='jitter', alpha=0.3)

```

An increase in alcohol does decrease the density which is something we 
found earlier. Lower density and higher alcohol do contribute to a better quality

### Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest? Were there any interesting or surprising interactions between features?

Alcohol and density correlate strongly and impact the quality of the wine. 
Alcohol has a lower density than water and hence an increase in alcohol will
reduce the density of the wine. However I cannot say quality of the wine is
related to the increase in alcohol or the fact the alcohol changes the texture
of the wine making.

The plots also strenghtened the relationship with citric acid having a positive
effect on quality and volatile acid having a negative effect on the quality.

Fixed acidity didn't impact quality as much as I expected it it. Citric acidity
is classed as as fixed acid, there are also other fixed acids such as tartaric 
and malic. These don't seem to have as great effect on quality as citric does.


## Final plots and summary
___

### Plot 1

```{r echo=FALSE, final_plot_volatile_acidity, message=FALSE, warning=FALSE}

ggplot(aes(x=alcohol,fill=q_factors), data=wine_dataset) +
  geom_histogram(bins=20) +
  xlab("Alcohol percentage in the wine (%)") +
  ylab("Count of number of wines with specific alcohol content") +
  ggtitle("Distribution of Alcohol % content, visible by Quality of wine") +
  theme(legend.title = element_text("Quality Factors"))
```

### Description for Plot 1

The distribution of alcohol is shown above. We can see that that very good 
quality wines have more alcohol contents.

### Plot 2

```{r echo=FALSE, message=FALSE, warning=FALSE}

create_quality_boxplot <- function(y.variable, y.title) {
  
  selected.column <- wine_dataset[,y.variable]
  
  bplot <- ggplot(aes_string(x='q_factors',y=y.variable, fill='q_factors'),
      data=wine_dataset) +
    geom_boxplot(outlier.shape = 4) +
    xlab("Wine quality rating") +
    ylab(y.title) +
    ylim(min(selected.column), quantile(selected.column, 0.99)) +
    theme(legend.position="none")
  return(bplot)
}



b_plot_alcohol <- create_quality_boxplot('alcohol', 'Alcohol by percentage (%)')
b_plot_acidity <- create_quality_boxplot('volatile.acidity', 'Volatile Acidity in g/dm^3')
b_plot_citric_acid <- create_quality_boxplot('citric.acid', 'Citric Acid in g/dm^3')
b_plot_sulphate <- create_quality_boxplot('sulphates', 'Potassium Sulphate - g/dm^3')


# arrange plots in a grid
grid.title <- "Box plots showing Quality of Wine for (1) Alcohol (%) (2) Volatile acidity ,(3) Citric acid, (4) Sulphate"
grid.arrange(b_plot_alcohol,b_plot_acidity,b_plot_citric_acid, b_plot_sulphate, top=grid.title)


```

### Description for Plot 2

Four boxplots are arranged in a grid showing different properties of the wine.
Each boxplot shows the distribution of data depending on rating the wine was 
given.

The median value for Alcohol for bad and ok wines stays roughly the same, 
however we notice an increase in alcohol for the best quality wines.

The best quality wines also exhibit the least acedic acid but contain the most
citric acid. 

There is also a gradual increase in sulphate, the sulphate contributes so 
sulfur dioxide which reduces the oxidation and microorganisms. 

### Plot 3

```{r echo=FALSE, message=FALSE, warning=FALSE}
# create the scatter plot

ggplot(aes(x=volatile.acidity, y=citric.acid), data=wine_dataset) +
  geom_point(aes(color=q_factors), position='jitter', alpha=0.6) +
  ylim(0,quantile(wine_dataset$citric.acid, quantile.level)) +
  xlab("Volatile Acidity in g/m^3") +
  ylab("Citric Acid in g/m^3") +
  ggtitle("Scatter plot of Citric acid against Volatile acidity, colors by Quality rating") +
  scale_color_brewer(type='qual', palette=1, guide = guide_legend(title = "Quality Rating"))
```

#### Description of Plot 3

Plot 3 is a scatter plot showing volatile acidity on the x axis and citric acid
on the y axis. We can see the colors of the wine quality ratings, the legend
on the right shows the mappings for bad, ok and great quality wines.

It shows that citric acid is found in good quality wines, higher volatile
acidity is shown in wines of poorer quality.


## Reflection
___

Out of all the variables studied, the strongest relationships were quite 
evident early on. Volatile acid, alcohol, citric acid sulphates seem to have
the greatest impact on the quality rating of the wine.

Volatile acidity can reduce the taste the aroma of the wine. Excess acetic acid
also puts the yeast under stress which ends up producing more volatile acid as
a result. 

Citric acid is used to add flavour to the wine, so this will make the wine 
taste nicer and hence get a better quality rating.

Alcohol was also found to increase the quality of the wine, it impacted the 
density which means the wine was less dense.

Sulphates ensure enough sulfur dioxide is present in the wine to not cause
too many miroorganisms to proliferate and to ensure oxidation does not happen
too early. Too much can be detected in in the flavour which is undesirable,
this is possibly why there was a sudden reduction in sulphate quantity in the
wine after 1.0 g/m^3.

There were some limitations in the dataset. I was interested to see the
different vineyards the grapes were grown from, particularly the climate in 
those countries.

If wine is produced in warmer climates means the grapes become fully ripe but
they contain less acidity. cooler climates tend to have more acidity. These
wines produces in warmer climates usually have acid added into the wine.

However I understand the manufacturers wanted company names removed from the 
dataset.

Overall this was an enjoyable experience, exploring the dataset was interesting
and a challenge. I had a lot of fun and it was a great way to go over some of
the R code I had learnt.

## References
___


[1] https://en.wikipedia.org/wiki/Fermentation_in_winemaking
[2] http://winefolly.com/tutorial/warm-climate-vs-cool-climate-wine/
[3] http://www.gencowinemakers.com/docs/Sulfur%20Dioxide.pdf



