---
title: "Lesson 7 "
output: html_notebook
---

```{r}
# clear variables from previous sessions
rm(list=ls())
```


## Third qualitative variable

### read in the data 
```{r}
pf = read.csv('D:/fb_data/pseudo_facebook.tsv', sep = '\t')
```

###

```{r}
ggplot(aes(x=gender, y=age), data=subset(pf, !is.na(gender))) +
  geom_boxplot() +
  # let's add the mean to the box plot too
  stat_summary(fun.y = mean, geom='point', shape=4, size=5, color='red')
```



```{r}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(gender))) + 
    geom_line(aes(color=gender), stat= 'summary', fun.y=median)
```


# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.
```{r}

# I think we're grouping by age and further grouping by
library(dplyr)
grouped_data <- group_by(subset(pf, !is.na(gender)), age, gender)
pf.fc_by_age_gender <- summarise(grouped_data, mean_friend_count=mean(friend_count), median_friend_count = median(friend_count), n=n())
pf.fc_by_age_gender

```

## using the chain command

```{r}
pf.fc_by_age_gender2 <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age,gender) %>%
  # summarise removes one layer of grouping when it runs
  # it removes the last group. so it will remove the gender layer
  summarise(mean_friend_count=mean(friend_count), median_friend_count = median(friend_count), n=n()) %>%
  ungroup() %>%
  arrange(age)
pf.fc_by_age_gender2

```


# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.


```{r}
ggplot(aes(x=age, y=median_friend_count), data = pf.fc_by_age_gender2) + 
  geom_line(aes(color=gender))
```

# going to convert from long to wide format, can use reshape2 package or tidr, example of long and wide
# long

# age   |   gender  |   median_friend_count |
# ======|===========|=======================|
# 13    |   male    |         254           |
# 13    |   female  |         180           |
# 14    |   male    |         63            |

# wide

# age   | male (median) | female (median) |
# ======|===============|=================|

```{r}
install.packages("tidyr")
```

```{r}
library(tidyr)

# spread function
# spread(data, key, value)

pf.fc_by_age_gender.wide1 <- spread(
  subset(pf.fc_by_age_gender2, select = c("gender", "age", "median_friend_count")),
  gender,median_friend_count)
pf.fc_by_age_gender.wide1


```

# using reshape2 to reshape the data

```{r}
install.packages('reshape2')
library(reshape2)
```

```{r}
# dcast - result to be a dataframe
# acast - result to be an array

# dcast(dataset, formula)
# formula = what you want to keep on the left, + sign to separate ~ gender, then the actual value of the thing you split, using value.var

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender2,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

pf.fc_by_age_gender.wide
```


## plot the ratio of females to males to see the number of friends the average fe

```{r}
pf.fc_by_age_gender.wide1$ratio = pf.fc_by_age_gender.wide1$female/pf.fc_by_age_gender.wide1$male
pf.fc_by_age_gender.wide1

ggplot(aes(x=age, y=ratio), data = pf.fc_by_age_gender.wide1) +
  geom_line() + 
  geom_hline(yintercept = 1, linetype = 2)
```


## third quantitative variable 

# tenure - how long you've been on fb
# let's group users by the year they joined

```{r}

pf_2 <- subset(pf, select = c("userid", "age", "tenure", "friend_count", "gender")) 
pf_2$year_joined <- floor((2014 - (pf_2$tenure / 365)))
pf_2

summary(pf_2$year_joined)

```


# using cut to create 4 bins across the dates
# 2004 - 2009
# 2009 - 2011
# 2011 - 2012
# 2012 - 2014

```{r}
new_cut <- cut(pf_2$year_joined, 4)
# this unfortunatley gives the following dates
# (2005,2007] (2007,2010] (2010,2012] (2012,2014
year_cuts <- c(2004, 2009, 2011, 2012, 2014)
new_cut <- cut(pf_2$year_joined, breaks = year_cuts)
levels(new_cut)

pf_2$year_joined.bucket <- new_cut
table(pf_2$year_joined.bucket, useNA = 'ifany')
#typeof(pf_2$year_joined)
```

## plot the median friend_count across the age factors

```{r}
ggplot(aes(x=age, y=friend_count), data = subset(pf_2, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = median)
  
```
## Grand Mean

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

```{r}
ggplot(aes(x=age, y=friend_count), data = subset(pf_2, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y=mean,linetype=2)
  
  
```

# find the number of friends each user has per day since starting facebook

```{r}

pf = read.csv('D:/fb_data/pseudo_facebook.tsv', sep = '\t')
pf_3 <- subset(pf,select = c("tenure", "friend_count"), tenure > 0)
pf_3$friend_rate <- pf_3$friend_count / pf_3$tenure
max(pf_3$friend_rate)

```
## create line graph of friendships intiated per day vs tenure
# you need age, tenure , friendships_initated, year_joined.bucket


```{r}
pf_4 <- subset(pf, select = c("age", "tenure","friendships_initiated"), tensure > 0)
pf_4$year_joined <- floor((2014 - (pf_4$tenure / 365)))

pf_4$year_joined.bucket <- cut(pf_4$year_joined, breaks = year_cuts)
pf_4$friendships_initiated_rate <- pf_4$friendships_initiated/pf_4$tenure

ggplot(aes(x=tenure, y=friendships_initiated_rate), data = pf_4) + 
  geom_line(aes(color=year_joined.bucket),stat='summary', fun.y='mean')
```


## using geom_smooth


```{r}
ggplot(aes(x=tenure, y=friendships_initiated_rate), data = pf_4) + 
  geom_smooth(aes(color=year_joined.bucket))
```

# Yogurt data


```{r}
yo <- read.csv('D:/RFiles/yogurt.csv')

# convert yo id to factor variable

yo$id <- factor(yo$id)
str(yo)
```

```{r}

qplot(x=price, data=yo, binwidth=10)

```


```{r}

unique(yo$price)
```

# now find the total counts of yogurt for each household purchase.... using the transform function
# transform allows you to create a dataframe, it can also combine variables in the dataframe

```{r}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
qplot(x=all.purchases, data=yo, binwidth=1)
```


```{r}
ggplot(aes(x = time, y=price), data=yo) + geom_point()
#ggplot(aes(x=time, ))
```

```{r}
levels(yo$id)
```

```{r}
set.seed(4230)

#sample from the factors, let#s sample 16 ids
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price), data= subset(yo, id %in% sample.ids)) + 
  facet_wrap(~ id) +
  geom_line() +
  geom_point(aes(size=all.purchases), pch=1)

```

## scatterplot matrices

# going to use the ggally library, an extension to 

```{r}
install.packages('GGally')
library(GGally)
```

```{r}
theme_set(theme_minimal(20))

set.seed(1836)

pf_subset <- pf[,c(2:15)]

# nrow function returns the number of rows present
# sample.int 

ggapirs(pf_subset[sample.int()])
```
---
title: "lesson_7"
output: html_document
---


# scatter plot matrix great when all your variables are numerical, but sometimes you get categorical variables in there so regular scatterplot isn't good enough


```{r}
install.packages('GGally')
```

```{r}
library(GGally)
library(ggplot2)
```

```{r}

pf <- read.csv('D:/RFiles/lesson3/pseudo_facebook.tsv', sep = '\t')
pf_subset <- pf[,c(2:15)]

theme_set(theme_minimal(20))
set.seed(1836)
# get the number of rows in the 
n_rows <- nrow(pf_subset)

# gets a sample of ints, you specify the max number of the variable and the same size, ensure sample size < max number or ensure to add replace=False

# generate 1000 samples from the dataset
random_sample <- sample.int(n_rows, 1000)
# dataset [rows,cols] (leave empty for all rows or cols)

ggpairs(pf_subset[random_sample,])

```

```{r}
library(reshape2)
```

# reading in genomic data - Heatmap
```{r}
nci <- read.table('D:/RFiles/nci.tsv')
# new column names are between 1 - 64
colnames(nci) <- c(1:64)

#the melt command takes in a dataset, some the column names are placed in the dataset as a variable
#  subject time age weight height
#1    John    1  33     90      2
#2    Mary    1  NA     NA      2

# molten = melt(x, id = c("subject", "time"))
# molten
#   subject time variable value
# 1    John    1      age    33
# 2    Mary    1      age    NA
# 3    John    1   weight    90
# 4    Mary    1   weight    NA
# 5    John    1   height     2
# 6    Mary    1   height     2

nci.long.samp <- melt(as.matrix(nci[1:200,]))
#rename
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

```

```{r}
ggplot(aes(y=gene, x=case, fill=value), data=nci.long.samp) + 
  geom_tile() +
  scale_fill_gradientn(colors=colorRampPalette(c('blue','red'))(100))
```














































