---
title: "Lesson_9_diamondsandprice_predictions"
output: html_document
---

```{r}
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')
```
```{r}
library(GGally)
library(scales)
#library(memisc)
library(lattice)
library(MASS)
library(car)
library(reshape)
library(plyr)
```

```{r}
rm(list=ls())

library(ggplot2)
```

```{r}

data(diamonds)
head(diamonds)

```


# Quiz 3 - Scatterplot Review

# Let's start by examining two variables in the data set.
# The scatterplot is a powerful tool to help you understand
# the relationship between two continuous variables.

# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond
# characteristics to help us figure out whether
# the price advertised for any given diamond is 
# reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.



```{r}
percentile <- 0.99
ggplot(aes(x=carat, y=price),data=diamonds) +
  geom_point( alpha=0.05, shape=21) +
  xlim(0, quantile(diamonds$carat, percentile)) +
  ylim(0, quantile(diamonds$price, percentile)) + 
  stat_smooth(method = 'lm')
```
the graph shows a non linear relationship, slightly exponential.
- the variance in price is much larger as the carat increases

# sample the diamond dataset, we're going to do a pairwise plot, it's going to be huge so sample the dataset



```{r}
set.seed(20022012)

# sample

diamond_sample <- diamonds[sample(),]

```