---
title: "Lesson_8"
output: html_document
---

loading in data 
```{r}
rm(list = ls())
library(ggplot2)
```


# Quiz 1

# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

```{r}
qplot(x=price, data=diamonds, fill=cut) +
  facet_wrap(~color) +
  scale_x_log10()
```

# Quiz 2

# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

```{r}
ggplot(aes(x=table, y=price),data=diamonds) +
  geom_point(aes(color=cut)) + 
  xlim(50,80)
```

# Quiz 3

# What is the typical table range for the majority of the diamonds  of idead cut?

```{r}
ideal_cut <- subset(diamonds, cut=='Ideal')
summary(ideal_cut$table)

```

# What is the typical table range for the majority of the diamonds  of premium cut?

```{r}
premium_cut <- subset(diamonds, cut=='Premium')
summary(premium_cut$table)

```

# Quiz 4

# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')



```{r}
new_diamonds <- transform(diamonds, volume = x*y*z)


quantile(new_diamonds$price,0.99)

# could have used filter aswell

ggplot(aes(x=volume, y=price), data=subset(new_diamonds,price < quantile(new_diamonds$price,0.99))) +
  geom_point(aes(color=clarity)) +
  scale_y_log10() +
  xlim(0,400)

```

# Quiz 5

# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

```{r}
pf <- read.csv('D:/RFiles/lesson3/pseudo_facebook.tsv', sep = '\t')

pf_2 <- transform(subset(pf, select=c("userid", "friend_count", "friendships_initiated")), prop_initiated=ifelse(friendships_initiated ==0, 0, friendships_initiated/friend_count))


```

# quiz 6

# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg


```{r}
pf_3 <- transform(subset(pf, select=c("userid", "friend_count", "friendships_initiated", "tenure")), prop_initiated=ifelse(friendships_initiated ==0, 0, friendships_initiated/friend_count))

pf_3 <- transform(pf_3, year_joined = floor(2014 - (tenure/365)))
pf_3 <- transform(pf_3, year_joined.bucket = cut(year_joined, c(2004,2009,2011,2012,2014)))

ggplot(aes(x=tenure, y=prop_initiated), data=pf_3) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median) + 
  geom_smooth()

```

# for the gorup with the largest proportion of friendships intiated, what is the group;s average proportion of friendships initiated

```{r}

newest_members <- subset(pf_3, year_joined.bucket == levels(pf_3$year_joined.bucket)[3])
summary(newest_members$friendships_initiated)
```


# quiz 10

# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

```{r}
head(diamonds)
```

```{r}

ggplot(aes(x=cut, y=price/carat), data=diamonds) +
  facet_wrap(~clarity) + 
  geom_point(aes(color=color))

ggsave("D:/RFiles/price_carat.png")

```


