---
title: 'Lesson 9 : Diamonds & Price Predictions'
output: html_document
---


```{r}
rm(list = ls())
```

```{r}
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('plyr')
install.packages('gridExtra')
```

```{r}
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
data("diamonds")
```


# scatterplot

```{}

```

# GGPairs 

```{r}
set.seed(20022012)


# sample 1000 elements of the diamonds dataset
# sample will by default sample the columns, so can;t just use the sample function on the dataset

diamond_sample <- diamonds[sample.int(nrow(diamonds),1000),]


# the lower and the upper parameters take a list, which also has a variable the variables can be
# continuous, combo, discrete, na
ggpairs(diamond_sample,
        lower = list(continuous = wrap("points", shape= I('.'))),
        upper = list(combo = wrap("box", outlier.shape = I('x')))
        )

```

# Quiz 8


# Create two histograms of the price variable
# and place them side by side on one output image.

# We've put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

# You can self-assess your work with the plots
# in the solution video.

# ALTER THE CODE BELOW THIS LINE
# ==============================================
```{r}
library(gridExtra)

plot1 <- qplot(x=price, data=diamonds, binwidth=100) + 
  ggtitle('Price')

plot2 <- qplot(x=price, data=diamonds, binwidth=0.01) +
  scale_x_log10()+
  ggtitle('Price (log10)')

grid.arrange(plot1, plot2)

```

The 2 peaks show the demand in diamonds for cheaper and the more expensive
- bimodal, 2 classes, rich and poor customers

# scatterplot transformation
```{r}


qplot(x=carat, y=price, data=diamonds) +
  #scale_y_log10() + 
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('price (log10) by carat')
```

# scatterplot transformation using cuberoot

```{r}
cuberoot_trans = function() trans_new('cuberoot',
          transform = function(x) x^(1/3),
          inverse = function(x) x^3)

# implicit return 
noodle = function(x) 2*x 



ggplot(aes(x=carat, y=price), data = diamonds) +
  geom_point(alpha=0.01, position="jitter", size=0.75) +
  scale_x_continuous(trans = 'cuberoot', limits=c(0.2,3), breaks=c(0.2,0.5,1,2,3)) +# i think you can even pass cuberoot_trans()
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('log10 price vs cuberoot carat')
```

## overplotting - when multiple points take on the same value, which is why we sometimes introduce jitter

```{r}
# table tells you each unique variable and the counts of each variable
table(diamonds$carat)

# let's gett he variable counts and sort 
# I've also seen decreasing = T
sort(table(diamonds$carat), decreasing = TRUE)

head(sort(table(diamonds$carat), decreasing = TRUE))

```


# Using color

```{r}
install.packages('RColorBrewer', dependencies = TRUE)
library(RColorBrewer)

```

```{r}
ggplot(aes(x=carat, y=price), data = diamonds) +
  geom_point(alpha=0.01, position="jitter", size=0.75, aes(color=clarity)) +
  scale_x_continuous(trans = 'cuberoot', limits=c(0.2,3), breaks=c(0.2,0.5,1,2,3)) +# i think you can even pass cuberoot_trans()
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('log10 price vs cuberoot carat')

```

```{r}
ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter', aes(color=clarity)) +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```


# clarity does explain a lot of the variance



# examining cut

```{r}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and cut')
```

# linear models


# lm(y~x) where y is the outcome variable and x is the explanatory variable

# in our study we'd apply the formula log(price)~carat^(1/3)

# I() wrapper, I stands for as is, it tells I to transform the variable BEFORE doing the regression

```{r}
model_1 <- lm(I(log(price))~I(carat^(1/3)), data=diamonds)

# you can add further variables to the model, to see what explains most of the variance

model_2 <- update(model_1, ~ . + carat) # how much does untransformed carat explain the variance
model_3 <- update(model_2, ~ . + cut) # how much does untransformed carat explain the variance
model_4 <- update(model_3, ~ . + color) # how much does untransformed carat explain the variance
model_5 <- update(model_4, ~ . + clarity) # how much does untransformed carat explain the variance

mtable(model_1,model_2,model_3, model_4, model_5)

```

# predict the model


# at 0.95 confidence interbal
```{r}
# when the model has been built, you can use it to predict, using the predict function
unknown <- data.frame(carat=0.6, cut='V.good')
model_estimate <- predict(model_2, newdata = unknown, interval = "prediction", level=.95)


model_estimate
```





